Global phase diagram and possible quantum spin liquid in the triangular $J-K-\Gamma$ model

### 2020-03-26 paper/paper.tex
20200326 12:43
Signed-off-by: wangshiphys <wangshiphys@gmail.com>

\begin{figure}
    \includegraphics[width=\columnwidth]{Fig1.pdf}
    \caption{\label{fig:ModelDefinition}(Color online) (a) Top view of the triangular lattice of the edge-sharing octahedron. (b) The orientation of the cubic $x$, $y$, $z$ axes with respect to the octahedron. The spin operators $S^x$, $S^y$ and $S^z$ are defined with respect to this reference frame. (c) Three types of the NN bonds on the triangular lattice, namely $\gamma=x, y, z$ colored red, green and blue, respectively. The different colors of the lattice sites label the four sublattices realizing the four-sublattice-transformation, see tha main text for detailed information. (d) The k-points in the first Brillouin zone corresponding to a 4 $\times$ 6 cluster with 4 lattice sites along the $x$-bond direction and 6 lattice sites along the $y$-bond direction.}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig3.pdf}
    \caption{\label{fig:StructureFactors}(Color online) SMSF for different model paramters. The green dashed lines marks the first BZ corresponding to the triangular lattice.}
\end{figure}
\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig4.pdf}
    \caption{\label{fig:Spectrum}(Color online) Spin excitation spectrum for different model parameters. The $A(\bm{k}, \omega)$ is draw along the $M-\Gamma-M$ path in the first BZ. For the 4 $\times$ 6 cluster we use, there are 24 nonequivalent k-points in the first BZ, and 7 k-points on the $M-\Gamma-M$ path, see Fig.~\ref{fig:ModelDefinition}(d)}
\end{figure}

When the off-diagonal $\Gamma$ term is included, the transforamtion is no longer valid. To identify the remaining phases, we calculate the SMSF and spin excitation spectrum for some representative model paramters, as shown in Fig.~\ref{fig:StructureFactors} and Fig.~\ref{fig:Spectrum}. For pure $\Gamma$ model, see Fig.~\ref{fig:StructureFactors}(a) and (b), the peak of the SMSF locates at the center of first Brillouin Zone (BZ) which is a typical characteristic of the FM order. Based on this observation, we propose the three blue areas with different shadings in the global phase diagram (see Fig.~\ref{fig:PhaseDiagram}) to be FM ordered phases (i.e., FM-A, FM-B, FM-C) and they different from each other in the direction of the ordered moment which will be discussed in the Sec.~\ref{sec:SectionV}. The ground states of the quantum model in these area are similar to their classical analogue except for the pure $\Gamma$ model. For $\Gamma=1$, the GS of the classical model is highly degenerate including FM ordered states and states with no long range order. However, for the quantum model, quantum flucations select the FM ordered states as GS out of an otherwise infinitely degenerate manifold of classical states. This selection of states among a continuum is the so called "order-from-disorder" mechanism previously emphasized by Villain and Henley in the context of the square lattice XY model \cite{PhysRevLett.62.2056} and by Oguchi \emph{et al.,} in the case of fcc antiferromagnet \cite{JPSJ.54.4494}.

In Fig.~\ref{fig:StructureFactors}(c)-(e), the SMSF peaked at the M-points indicating that the stripe ordered GS of the classical model persist in the quantum limit. It is worth to note that even the GS of the classical model near the AF Kitaev point was identified to be nematic ordered which is energetically degenerate with the stripe state, our ED results favor the idea that quantum flucations would select a stripe state. The four stripe phases designated as Stripe-A, Stripe-B, Stripe-C, Stripe-D in Fig.~\ref{fig:PhaseDiagram}(a) different from each other only in the moment direction. Apart from these single-$\mathbf{Q}$ phases, we also identify a multi-$\mathbf{Q}$ phase, see the green area in Fig.~\ref{fig:PhaseDiagram}(a). For $J$, $K$, $\Gamma$ in this area, the SMSF has high intensities at both the center and M-points of the first BZ and Fig.~\ref{fig:StructureFactors}(f) is a typical SMSF profile of this phase.

>>>>

\begin{figure}
    \includegraphics[width=\columnwidth]{Fig1.pdf}
    \caption{\label{fig:ModelDefinition}(Color online) (a) Top view of the triangular lattice of the edge-sharing octahedron. (b) The orientation of the cubic $x$, $y$, $z$ axes with respect to the octahedron. The spin operators $S^x$, $S^y$ and $S^z$ are defined with respect to this reference frame. (c) Three types of the NN bonds on the triangular lattice, namely $\gamma=x, y, z$ colored red, green and blue, respectively. The different colors of the lattice sites label the four sublattices realizing the four-sublattice-transformation, see tha main text for detailed information.}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig3.pdf}
    \caption{\label{fig:StructureFactors}(Color online) SMSF for different model paramters. The green dashed lines marks the first BZ corresponding to the triangular lattice.}
\end{figure}

When the off-diagonal $\Gamma$ term is included, the transforamtion is no longer valid. To identify the remaining phases, we calculate the SMSF for some representative model paramters, as shown in Fig.~\ref{fig:StructureFactors}. For pure $\Gamma$ model, see Fig.~\ref{fig:StructureFactors}(a) and (b), the peak of the SMSF locates at the center of first Brillouin Zone (BZ) which is a typical characteristic of the FM order. Based on this observation, we propose the three blue areas with different shadings in the global phase diagram (see Fig.~\ref{fig:PhaseDiagram}) to be FM ordered phases (i.e., FM-A, FM-B, FM-C) and they different from each other in the direction of the ordered moment which will be discussed in the Sec.~\ref{sec:SectionV}. The ground states of the quantum model in these area are similar to their classical analogue except for the pure $\Gamma$ model. For $\Gamma=1$, the GS of the classical model is highly degenerate including FM ordered states and states with no long range order. However, for the quantum model, quantum flucations select the FM ordered states as GS out of an otherwise infinitely degenerate manifold of classical states. This selection of states among a continuum is the so called "order-from-disorder" mechanism previously emphasized by Villain and Henley in the context of the square lattice XY model \cite{PhysRevLett.62.2056} and by Oguchi \emph{et al.,} in the case of fcc antiferromagnet \cite{JPSJ.54.4494}. Similarly, the degeneracy for the classical model at $\Gamma=-1$ is lifted by quantum flucations and the FM ordered state is selected as the GS.

For these areas marked as Stripe-A, Stripe-B, Stripe-C and Stripe-D in the global phase diagram, the representative SMSF profiles are shown in Fig.~\ref{fig:StructureFactors}(c)-(e). For all of these model parameters, the SMSF peaked at the midpoints of the first BZ boundary (i.e., M-points) indicating that the stripe ordered GS of the classical model persist in the quantum limit. It is worth to note that even the GS of the classical model near the AF Kitaev point was identified to be nematic ordered which is energetically degenerate with the stripe state, our ED results favor the idea that quantum flucations would select a stripe state. The four stripe phases designated as Stripe-A, Stripe-B, Stripe-C, Stripe-D in Fig.~\ref{fig:PhaseDiagram}(a) different from each other only in the moment direction.

\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig4.pdf}
    \caption{\label{fig:Spectrum}(Color online) (a) The k-points in the first Brillouin zone corresponding to a 4 $\times$ 6 cluster with 4 lattice sites along the $x$-bond direction and 6 lattice sites along the $y$-bond direction. (b) Spin excitation spectrum for $\alpha=0.05\pi,\beta=0.10\pi$. The $A(\bm{k}, \omega)$ is draw along the path shown in (a).}
\end{figure}

Apart from these single-$\mathbf{Q}$ phases, we also identify a multi-$\mathbf{Q}$ phase, see the green area in Fig.~\ref{fig:PhaseDiagram}(a). For $J$, $K$, $\Gamma$ in this area, the SMSF has high intensities at both the center and M-points of the first BZ. Fig.~\ref{fig:StructureFactors}(f) is a typical SMSF profile of this phase. To have further understanding of the characteristics of this phase, we also calculate the spin excitation spectrum and a typical $A(\mathbf{k}, \omega)$ profile is shown in Fig~\ref{fig:Spectrum}(b). The high intensities at both the center and M-points of the first BZ is consistent with SMSF data. Moreover, there are continuum spectrum at high energy which is a characteristic of QSL. However, the $A(\mathbf{k}, \omega)$ does not have intensities at $\omega = 0$ which indicates the system described by the corresponding model Hamiltonian is gapped. Combined with all of the above results, we propose the GS of quantum $J-K-\Gamma$ model at the green area is a gapped QSL.
